---
title: "Donald Kane"
output: github_document
---

Find the correlation between energy consumption, production and expenditures and use scatterplot with a regression line to show the relationship. (scatterplot)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Use U.S. maps to show the consumption and expenditures of coal, petroleum, etc. Find each state’s consumption to GDP correlation. (US map plots, scatterplots, correlation)

```{r}
library(readr)
library(usmap)
library(ggplot2)
library(tidyverse)
library(patchwork)
```

```{r}
Energy_Census_and_Economic_Data_US_2010_2014 <- read_csv("Energy Census and Economic Data US 2010-2014.csv")
Energy_Census_and_Economic_Data_US_2010_2014 <- Energy_Census_and_Economic_Data_US_2010_2014[1:51,]
View(Energy_Census_and_Economic_Data_US_2010_2014)
```


``` {r}
data <- Energy_Census_and_Economic_Data_US_2010_2014 %>%
  filter(State != "District of Columbia") %>%
  filter(State != "United States")
colnames(data)[2] <- "state"

pcc1 <- plot_usmap(data = data, values = "TotalC10-11", color = "black") + 
  scale_fill_continuous(name = "Consumption Change (%)", label = scales::comma, low = "white", high = "red") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Consumption 2010-2011"
  )

pcc2 <- plot_usmap(data = data, values = "TotalC11-12", color = "black") + 
  scale_fill_continuous(name = "Consumption Change (%)", label = scales::comma, low = "white", high = "red") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Consumption 2011-2012"
  )

pcc3 <- plot_usmap(data = data, values = "TotalC12-13", color = "black") + 
  scale_fill_continuous(name = "Consumption Change (%)", label = scales::comma, low = "white", high = "red") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Consumption 2012-2013"
  )

pcc4 <- plot_usmap(data = data, values = "TotalC13-14", color = "black") + 
  scale_fill_continuous(name = "Consumption Change (%)", label = scales::comma, low = "white", high = "red") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Consumption 2013-2014"
  )

pcc1 + pcc2 + pcc3 + pcc4
```

```{r}
pcp1 <- plot_usmap(data = data, values = "TotalP10-11", color = "black") + 
  scale_fill_continuous(name = "Production Change (%)", label = scales::comma, low = "white", high = "blue") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Production 2010-2011"
  )

pcp2 <- plot_usmap(data = data, values = "TotalP11-12", color = "black") + 
  scale_fill_continuous(name = "Production Change (%)", label = scales::comma, low = "white", high = "blue") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Production 2011-2012"
  )

pcp3 <- plot_usmap(data = data, values = "TotalP12-13", color = "black") + 
  scale_fill_continuous(name = "Production Change (%)", label = scales::comma, low = "white", high = "blue") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Production 2012-2013"
  )

pcp4 <- plot_usmap(data = data, values = "TotalP13-14", color = "black") + 
  scale_fill_continuous(name = "Production Change (%)", label = scales::comma, low = "white", high = "blue") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Production 2013-2014"
  )
pcp1 + pcp2 + pcp3 + pcp4
```

```{r}
pcpr1 <- plot_usmap(data = data, values = "TotalPrice10-11", color = "black") + 
  scale_fill_continuous(name = "Price Change (%)", label = scales::comma, low = "white", high = "green") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Price 2010-2011"
  )

pcpr2 <- plot_usmap(data = data, values = "TotalPrice11-12", color = "black") + 
  scale_fill_continuous(name = "Price Change (%)", label = scales::comma, low = "white", high = "green") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Price 2011-2012"
  )

pcpr3 <- plot_usmap(data = data, values = "TotalPrice12-13", color = "black") + 
  scale_fill_continuous(name = "Price Change (%)", label = scales::comma, low = "white", high = "green") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Price 2012-2013"
  )

pcpr4 <- plot_usmap(data = data, values = "TotalPrice13-14", color = "black") + 
  scale_fill_continuous(name = "Price Change (%)", label = scales::comma, low = "white", high = "green") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Price 2013-2014"
  )

pcpr1 + pcpr2 + pcpr3 + pcpr4
```

```{r}
pce1 <- plot_usmap(data = data, values = "TotalE10-11", color = "black") + 
  scale_fill_continuous(name = "Expenditure Change (%)", label = scales::comma, low = "white", high = "orange") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Expenditure 2010-2011"
  )

pce2 <- plot_usmap(data = data, values = "TotalE11-12", color = "black") + 
  scale_fill_continuous(name = "Expenditure Change (%)", label = scales::comma, low = "white", high = "orange") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Expenditure 2011-2012"
  )

pce3 <- plot_usmap(data = data, values = "TotalE12-13", color = "black") + 
  scale_fill_continuous(name = "Expenditure Change (%)", label = scales::comma, low = "white", high = "orange") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Expenditure 2012-2013"
  )

pce4 <- plot_usmap(data = data, values = "TotalE13-14", color = "black") + 
  scale_fill_continuous(name = "Expenditure Change (%)", label = scales::comma, low = "white", high = "orange") + 
  theme(legend.position = "right") +
  labs(
    title = "Percent Change in Total Energy Expenditure 2013-2014"
  )

pce1 + pce2 + pce3 + pce4
```

## Show distribution of energy consumption (bar graph)

```{r}
c1 <- data %>%
  ggplot(aes(`TotalC10-11`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalC10-11`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "red") + 
  theme_classic() + 
  labs(
    title = "2010 - 2011",
    x = "Energy Consumption in Billion BTU"
  )

c2 <- data %>%
  ggplot(aes(`TotalC11-12`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_density(alpha = 0.5, fill = "red") +
  geom_vline(aes(xintercept = mean(`TotalC11-12`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  theme_classic() + 
  labs(
    title = "2011 - 2012",
    x = "Energy Consumption in Billion BTU"
  )

c3 <- data %>%
  ggplot(aes(`TotalC12-13`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalC12-13`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "red") +
  theme_classic() + 
  labs(
    title = "2012 - 2013",
    x = "Energy Consumption in Billion BTU"
  )

c4 <- data %>%
  ggplot(aes(`TotalC13-14`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalC13-14`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "red") +
  theme_classic() + 
  labs(
    title = "2013 - 2014",
    x = "Energy Consumption in Billion BTU"
  )

c1 + c2 + c3 + c4
```

```{r}
pr1 <- data %>%
  ggplot(aes(`TotalP10-11`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 2.5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalP10-11`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "blue") + 
  theme_classic() + 
  labs(
    title = "2010 - 2011",
    x = "Energy Production in Billion BTU"
  )

pr2 <- data %>%
  ggplot(aes(`TotalP11-12`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 2.5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalP11-12`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "blue") +
  theme_classic() + 
  labs(
    title = "2011 - 2012",
    x = "Energy Production in Billion BTU"
  )

pr3 <- data %>%
  ggplot(aes(`TotalP12-13`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 2.5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalP12-13`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "blue") +
  theme_classic() + 
  labs(
    title = "2012 - 2013",
    x = "Energy Production in Billion BTU"
  )

pr4 <- data %>%
  ggplot(aes(`TotalP13-14`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 2.5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalP13-14`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "blue") +
  theme_classic() + 
  labs(
    title = "2013 - 2014",
    x = "Energy Production in Billion BTU"
  )

pr1 + pr2 + pr3 + pr4
```

```{r}
p1 <- data %>%
  ggplot(aes(`TotalPrice10-11`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalPrice10-11`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "green") + 
  theme_classic() + 
  labs(
    title = "2010 - 2011",
    x = "Average Price in Millions"
  )

p2 <- data %>%
  ggplot(aes(`TotalPrice11-12`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalPrice11-12`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "green") +
  theme_classic() + 
  labs(
    title = "2011 - 2012",
    x = "Average Price in Millions"
  )

p3 <- data %>%
  ggplot(aes(`TotalPrice12-13`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalPrice12-13`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "green") +
  theme_classic() + 
  labs(
    title = "2012 - 2013",
    x = "Average Price in Millions"
  )

p4 <- data %>%
  ggplot(aes(`TotalPrice13-14`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .5, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalPrice13-14`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "green") +
  theme_classic() + 
  labs(
    title = "2013 - 2014",
    x = "Average Price in Millions"
  )

p1 + p2 + p3 + p4
```

```{r}
e1 <- data %>%
  ggplot(aes(`TotalE10-11`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .995, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalE10-11`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "orange") + 
  theme_classic() + 
  labs(
    title = "2010 - 2011",
    x = "Total Energy Expenditures in Million USD"
  )

e2 <- data %>%
  ggplot(aes(`TotalE11-12`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .995, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalE11-12`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "orange") +
  theme_classic() + 
  labs(
    title = "2011 - 2012",
    x = "Total Energy Expenditures in Million USD"
  )

e3 <- data %>%
  ggplot(aes(`TotalE12-13`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .995, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalE12-13`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "orange") +
  theme_classic() + 
  labs(
    title = "2012 - 2013",
    x = "Total Energy Expenditures in Million USD"
  )

e4 <- data %>%
  ggplot(aes(`TotalE13-14`)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = .995, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(`TotalE13-14`, na.rm = T)),
          colour = "black", linetype ="longdash", size = .8) +
  geom_density(alpha = 0.5, fill = "orange") +
  theme_classic() + 
  labs(
    title = "2013 - 2014",
    x = "Total Energy Expenditures in Million USD"
  )

e1 + e2 + e3 + e4
```

```{r}
coalC <- data %>%
  select("state", "CoalC2010", "CoalC2011", "CoalC2012", "CoalC2013", "CoalC2014") %>%
  pivot_longer(c("CoalC2010", "CoalC2011", "CoalC2012", "CoalC2013", "CoalC2014"), names_to = "Year", values_to = "CoalConsumption")
coalC$Year <- gsub("CoalC", "", as.character(coalC$Year))
coalC

sc <- coalC %>%
  ggplot(subset(coalC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = CoalConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "Coal Consumption 2010 - 2014",
       x = "Year",
       y = "Coal Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(coalC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

``` {r}
bioC <- data %>%
  select("state", "BiomassC2010", "BiomassC2011", "BiomassC2012", "BiomassC2013", "BiomassC2014") %>%
  pivot_longer(c("BiomassC2010", "BiomassC2011", "BiomassC2012", "BiomassC2013", "BiomassC2014"), names_to = "Year", values_to = "BiomassConsumption")
bioC$Year <- gsub("BiomassC", "", as.character(bioC$Year))
bioC

sb <- bioC %>%
  ggplot(subset(bioC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = BiomassConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "BioMass Consumption 2010 - 2014",
       x = "Year",
       y = "BioMass Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(bioC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
elecC <- data %>%
  select("state", "ElecC2010", "ElecC2011", "ElecC2012", "ElecC2013", "ElecC2014") %>%
  pivot_longer(c("ElecC2010", "ElecC2011", "ElecC2012", "ElecC2013", "ElecC2014"), names_to = "Year", values_to = "ElecConsumption")
elecC$Year <- gsub("ElecC", "", as.character(elecC$Year))
elecC

se <- elecC %>%
  ggplot(subset(elecC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = ElecConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "Electricity Consumption 2010 - 2014",
       x = "Year",
       y = "Electricity Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(elecC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
fossC <- data %>%
  select("state", "FossFuelC2010", "FossFuelC2011", "FossFuelC2012", "FossFuelC2013", "FossFuelC2014") %>%
  pivot_longer(c("FossFuelC2010", "FossFuelC2011", "FossFuelC2012", "FossFuelC2013", "FossFuelC2014"), names_to = "Year", values_to = "FossFConsumption")
fossC$Year <- gsub("FossFuelC", "", as.character(fossC$Year))
fossC

sf <- fossC %>%
  ggplot(subset(fossC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = FossFConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "Fossil Fuel Consumption 2010 - 2014",
       x = "Year",
       y = "Fossil Fuel Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(fossC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
sc + sb + se + sf
```

```{r}
geoC <- data %>%
  select("state", "GeoC2010", "GeoC2011", "GeoC2012", "GeoC2013", "GeoC2014") %>%
  pivot_longer(c("GeoC2010", "GeoC2011", "GeoC2012", "GeoC2013", "GeoC2014"), names_to = "Year", values_to = "GeoConsumption")
geoC$Year <- gsub("GeoC", "", as.character(geoC$Year))
geoC

sg <- geoC %>%
  ggplot(subset(geoC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = GeoConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "Geothermal Consumption 2010 - 2014",
       x = "Year",
       y = "Geothermal Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(geoC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
hydroC <- data %>%
  select("state", "HydroC2010", "HydroC2011", "HydroC2012", "HydroC2013", "HydroC2014") %>%
  pivot_longer(c("HydroC2010", "HydroC2011", "HydroC2012", "HydroC2013", "HydroC2014"), names_to = "Year", values_to = "HydroConsumption")
hydroC$Year <- gsub("HydroC", "", as.character(hydroC$Year))
hydroC

sh <- hydroC %>%
  ggplot(subset(hydroC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = HydroConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "Hydro Consumption 2010 - 2014",
       x = "Year",
       y = "Hydro Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(hydroC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
natC <- data %>%
  select("state", "NatGasC2010", "NatGasC2011", "NatGasC2012", "NatGasC2013", "NatGasC2014") %>%
  pivot_longer(c("NatGasC2010", "NatGasC2011", "NatGasC2012", "NatGasC2013", "NatGasC2014"), names_to = "Year", values_to = "NatGasConsumption")
natC$Year <- gsub("NatGasC", "", as.character(natC$Year))
natC

sng <- natC %>%
  ggplot(subset(natC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = NatGasConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "Natural Gas Consumption 2010 - 2014",
       x = "Year",
       y = "Natural Gas Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(natC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
lpgC <- data %>%
  select("state", "LPGC2010", "LPGC2011", "LPGC2012", "LPGC2013", "LPGC2014") %>%
  pivot_longer(c("LPGC2010", "LPGC2011", "LPGC2012", "LPGC2013", "LPGC2014"), names_to = "Year", values_to = "LPGConsumption")
lpgC$Year <- gsub("LPGC", "", as.character(lpgC$Year))
lpgC

slpg <- lpgC %>%
  ggplot(subset(lpgC, !(state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas"))), mapping = aes(x = Year,y = LPGConsumption, group = state)) +
  geom_line(alpha = 0.5, size = 0.1) +
  labs(title = "LPG Consumption 2010 - 2014",
       x = "Year",
       y = "LPG Consumption",
       color="state") +
  theme_classic() +
  geom_line(data = subset(lpgC, state %in% c("Tennessee", "New Jersey", "New York", "California", "Texas")),aes(color = state), size = 0.5)
```

```{r}
sh + sg + sng + slpg
```
