---
title: "gdp"
author: "Anubha Nagar"
date: "28/11/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("hrbrthemes")
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(hrbrthemes)
library(ggplot2)
```

```{r}
data <- read_csv("Energy Census and Economic Data US 2010-2014.csv")
```

```{r}
head(data)
str(data)
```

```{r}
data=na.omit(data)
```

```{r}
#Filter data for gdp
#formulate table 
data<-data%>%
  select("StateCodes","State","Region","Division","Coast", "GDP2010", "GDP2011", "GDP2012", "GDP2014")%>%
  pivot_longer(c("GDP2010", "GDP2011", "GDP2012", "GDP2014"), names_to = "Year", values_to = "gdp")
#Clean Year BiomassC
data$Year<-gsub("GDP","",as.character(data$Year))
#Display the outcome
data
```


# understanding the increase in gdp over time in the US 

```{r}
overall_gdp <- data%>% 
  group_by(Year)%>%
  summarize(sum_gdp = sum(gdp))
overall_gdp
```
plotting the increase in gdp 

```{r}
ggplot(overall_gdp,aes(x = Year, y = format(sum_gdp, scientific = FALSE)))+ geom_point()+
geom_line( color="grey") +
    geom_point(shape=21, color="black", fill="#69b3a2", size=6) +
    theme_ipsum()+ ggtitle("Change in GDP over the years 2010- 2014") +labs(y= "", x = "GDP")
```

Yearly plot for top 5 state with highest gdp 

```{r}
top5_states <- data%>%
  group_by(State)%>%
  summarize(GDP_total = sum(gdp))%>%
  arrange(desc(GDP_total))%>%
  slice(1:5)
```

California, Texas, NY, Florida and Illinois has the highest cumulative GDP (2010- 2014)

Now lets plot to understand the trend in gdp for each of these cities.

```{r}
plot_cali <- data%>%
  filter(State == "California"| State == "Texas"|State == "New York"| State == "Illinois"|State == "Florida")
  ggplot(plot_cali, aes(x = Year, y = gdp))+ geom_point(aes(col = State))+geom_line(aes(group = State))
```

#Lets understand if the GDP of the state is above the mean gdp of US
for 2010

```{r}
mean_gdp <- data%>%
  filter(Year== 2013)%>%
  summarize(mean = (sum(gdp)/51))
mean_gdp
```

```{r}
data <- data%>%
  mutate(diff_from_mean = gdp - 314534.1)
```

```{r}
ggplot(data, aes(x = reorder(StateCodes,diff_from_mean), y = diff_from_mean)) +
  geom_bar(stat = "identity",
           show.legend = FALSE, fill = "lightblue") +
  xlab("Group") +
  ylab("Value") +
  coord_flip() + theme_grey(base_size = 5)+ ggtitle("2013 Analysis of GDP above/below the mean")
```


Now lets analyze if the disparity in GDP for the different regions of US

```{r}
nsew <- data%>%
  group_by(Region)%>%
  summarize(mean_gdp_region = mean(gdp))
nsew
```

plot bar plot to see the same 
northern states are doing the best in terms of gdp 


```{r}
ggplot(nsew, aes(x = Region, y = mean_gdp_region))+ geom_bar(stat = "identity", fill = "Purple", alpha = 0.4) + ylab("Mean GDP for each region")
```

which state has the greatest gdp increase?

```{r}
data2010 <- data%>%
  filter(Year == 2010)%>%
  select(StateCodes,gdp)
data2014 <- data%>%
  filter(Year == 2014)%>%
  select(StateCodes,gdp)
```

```{r}
df <- merge(x=data2010,y=data2014,by="StateCodes",all=TRUE)
df <- df%>%
  mutate(diff = gdp.y- gdp.x)%>%
  arrange(desc(diff))%>%
  slice(1:5)
df
```
Here we see the same states have the greatest increase in gdp as well as the mean gdp 

```{r}
ggplot(df, aes(x = StateCodes, y = format(diff), scientific = FALSE))+ geom_bar(stat = "identity", fill = "lightblue")+ ylab("GDP chnage between 2014 and 2010")
```


